---
title: "STRchiveLiteratureAutomation"
author: "Laurel Hiatt"
date: "2023-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(biomaRt)
library(rentrez)
library(dplyr)
library(easyPubMed)
library(stringr)
```

## Automating STRchive Literature Management

The goal for this Markdown is to be a repository of code that assists with 
literature management for the living database, STRchive.

Running tasks:
- Plot number of papers by year of first publication for sanity check
### Defining the Gene List
```{r genelist}
gene_list <- c("AFF2", "FMR2", "AFF3", "AR", "ARX", "ARX", "ATN1", "ATXN1", "ATXN10", "ATXN2",
               "ATXN3", "ATXN7", "ATXN8OS", "ATXN8", "BEAN1", "C9orf72", "CACNA1A", "CBL2", "COMP",
               "DAB1", "DIP2B", "DMD", "DMPK", "FMR1", "FOXL2", "FXN", "GIPC1", "GLS", "HOXA13",
               "HOXA13", "HOXA13", "HOXD13", "HTT", "JPH3", "LRP12", "MARCH6", "NIPA1", "NOP56",
               "NOTCH2NLC", "NUTM2B-AS1", "PABPN1", "PHOX2B", "POLG", "PPP2R2B", "PRDM12", "RAPGEF2",
               "RFC1", "RILPL1", "RUNX2", "SAMD12", "SOX3", "STARD7", "TBP", "TBX1", "TCF4", "TNRC6A",
               "XYLT1", "YEATS2", "ZIC2", "ZIC3", "ZNF713", "ZNF9", "CNBP1", "CSTB", "EIF4A3", "PRNP",
               "VWA1", "ABCD3", "FGF14")
```

### Getting gene synonyms from biomaRt based on gene list

```{r biomaRt}
# Specify the dataset to use (in this case, HGNC in human, using GRCh38 assembly)
mart <- useMart("ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", 
                host = "https://www.ensembl.org")

# Retrieve the HGNC symbol and Gene Synonym for the genes in the list
gene_info <- getBM(attributes = c("hgnc_symbol", "external_synonym"),
                   filters = "hgnc_symbol",
                   values = gene_list,
                   mart = mart)

#if I want to copy this whole thing
#clipr::write_clip(gene_info)

# Replace missing values in 'external_synonym' with 'hgnc_symbol'
# This is because some gene_names have no synonyms, so it's easier to group this way
gene_info$external_synonym[is.na(gene_info$external_synonym) | gene_info$external_synonym == ""] <- gene_info$hgnc_symbol[is.na(gene_info$external_synonym) | gene_info$external_synonym == ""]

```


### Get All Article Data for Each Locus in XML Format
```{r article data}
perform_pubmed_query <- function(gene_info) {
  #set up publications to be filled out
  publications <- list()

  #I have to do the queries separately or it breaks things
  for (i in 1:nrow(gene_info)) {
    gene_name <- gene_info$hgnc_symbol[i]
    gene_synonyms <- unlist(strsplit(gene_info$external_synonym[i], ", "))

    #basically, I want to group all gene synonyms with the root gene name
    search_terms <- c(gene_name, gene_synonyms)
    search_terms <- unique(unlist(search_terms))

    query <- paste0('("repeat expansion"[Title/Abstract] OR "tandem repeat"[
                    Title/Abstract] OR "repeat expansions"[Title/Abstract] OR
                    "tandem repeats"[Title/Abstract]) AND (', paste(search_terms,
                    collapse = '[Title/Abstract] OR '), '[Title/Abstract]) AND "English"[Language
                    ] AND ("disease"[Title/Abstract] OR "disorder"[Title/Abstract] OR "
                    diseases"[Title/Abstract] OR "disorders"[Title/Abstract] OR "
                    syndrome"[Title/Abstract] OR "syndromes"[Title/Abstract]) AND "
                    journal article"[Publication Type]) NOT "review"[Publication Type])')


    # Remove unnecessary slashes from search terms
    query <- gsub("\"", "", query, fixed = TRUE)
    
    search_results <- entrez_search(db = "pubmed", term = query)
    if (search_results$count > 0) {
      article_ids <- entrez_search(db = "pubmed", term = query)$ids
      articles <- entrez_fetch(db = "pubmed", id = article_ids, rettype = "xml")
      publications[[gene_name]] <- articles
    } else {
      publications[[gene_name]] <- NULL
    }
  }

  return(publications)
}

# Apply function
pubmed_results <- perform_pubmed_query(gene_info)

# Printing the PubMed query results for each gene
# This is pretty ugly but it's here to build on if necessary
for (gene_name in names(pubmed_results)) {
  print(paste("Gene:", gene_name))
  print(pubmed_results[[gene_name]])
}

```

### Geting Only the PMIDs for Each Locus
```{r PMIDs}
get_pmids <- function(gene_info) {
  pmid_list <- list()

  for (i in 1:nrow(gene_info)) {
    gene_name <- gene_info$hgnc_symbol[i]
    gene_synonyms <- unlist(strsplit(gene_info$external_synonym[i], ", "))

    search_terms <- c(gene_name, gene_synonyms)
    search_terms <- unique(unlist(search_terms))
    query <- paste0('("repeat expansion"[Title/Abstract] OR "tandem repeat"[
                    Title/Abstract] OR "repeat expansions"[Title/Abstract] OR
                    "tandem repeats"[Title/Abstract]) AND (', paste(search_terms,
                    collapse = '[Title/Abstract] OR '), '[Title/Abstract]) AND "English"[Language
                    ] AND ("disease"[Title/Abstract] OR "disorder"[Title/Abstract] OR "
                    diseases"[Title/Abstract] OR "disorders"[Title/Abstract] OR "
                    syndrome"[Title/Abstract] OR "syndromes"[Title/Abstract]) AND "
                    journal article"[Publication Type]) NOT "review"[Publication Type])')

    query <- gsub("\"", "", query, fixed = TRUE)
    search_results <- entrez_search(db = "pubmed", term = query)

    if (search_results$count > 0) {
      article_ids <- entrez_search(db = "pubmed", term = query, retmax = 10000)$ids
      pmid_list[[gene_name]] <- article_ids
    } else {
      pmid_list[[gene_name]] <- NULL
    }
  }

  return(pmid_list)
}

# Apply the function
pmid_results <- get_pmids(gene_info)

# Printing the retrieved PMIDs for each gene (much prettier than XMLs)
for (gene_name in names(pmid_results)) {
  print(paste("Gene:", gene_name))
  print(pmid_results[[gene_name]])
}

```


### Getting Earliest Pubdates and Titles for Each Locus (NEEDS TO BE UPDATED)
```{r df}
# Initialize an empty data frame
earliest_pub_dates_df <- data.frame(
  Gene = character(0),
  Earliest_Pub_Date = numeric(0),
  NumPub = numeric(0),
  Titles = I(list())
)

for (i in 1:length(pubmed_results)) {
  gene <- names(pubmed_results)[i]
  print(gene)
  # Extract publication years for each gene using custom_grep
  publication_years <- custom_grep(pubmed_results[[i]], "PubDate", "char")
  publication_years <- str_extract(publication_years, "(?<=<Year>)\\d{4}(?=</Year>)")

  # Convert years to numeric format
  publication_years <- as.numeric(publication_years)

  # Find the minimum publication year
  min_year <- min(publication_years, na.rm = TRUE)

  # Extract titles
  titles <- custom_grep(pubmed_results[[i]], "ArticleTitle", "char")

  # Count the number of titles
  titles_count <- length(titles)

  # Combine titles into a single string with index numbers
  titles_string <- paste0(seq_along(titles), ". ", titles, collapse = ", ")

  # Add values to the data frame
  row_index <- nrow(earliest_pub_dates_df) + 1
  earliest_pub_dates_df[row_index, ] <- list(gene, min_year, titles_count, I(titles_string))
}

```

### Making a plot based on the earliest dates

```{r plot}
earliest_pub_dates_df <- read.table("/Users/quinlan/Documents/Git/STRchive/scripts/earliest_pub_dates_only.tsv", header = TRUE, sep = "\t")


# Assuming 'pubs_and_dates' is your dataframe

library(ggplot2)
install.packages("ggrepel")
library(ggrepel)

ggplot(pubs_and_dates, aes(x = Earliest_Pub_Date, y = NumPub, label = Gene)) +
  geom_point(size = 3) +
  geom_text_repel(size = 3, nudge_x = 0.1, nudge_y = 0.2, angle = 90) +  # Adjust position of labels and rotate by 90 degrees
  labs(x = "Earliest Publication Date", y = "Number of Publications") +
  ggtitle("Earliest Publication Date vs Number of Publications") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14),
        axis.text = element_text(size = 10))


```
